Dominance Regions for Imaginary $\bfg$-vectors

### 2020-06-11 dominance_regions.tex
Proof of Lemma 3.5

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2j}$ joined to $t_{2j+1}$ by an edge labelled $1$ and $t_{2j-1}$ joined to $t_{2j}$ by an edge labelled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

@@

  Define \emph{two-parameter Chebyshev polynomials} $u_{i,\varepsilon}$ for $i\in\ZZ$ and $\varepsilon\in\{\pm\}$ recursively by
  \[u_{0,\varepsilon}=0,\quad u_{1,\varepsilon}=1,\quad u_{i+1,\varepsilon}=\begin{cases} bu_{i,-\varepsilon}-u_{i-1,\varepsilon} & \text{if $\varepsilon=+$;}\\ cu_{i,-\varepsilon}-u_{i-1,\varepsilon} & \text{if $\varepsilon=-$.} \end{cases}\]
  Observe that, by an easy induction, $u_{-i,\varepsilon}=-u_{i,\varepsilon}$.

  The recursion can be written more succinctly as $u_{i+1,\varepsilon}=u_{2,\varepsilon}u_{i,-\varepsilon}-u_{i-1,\varepsilon}$ with the additional initial condition $u_{2,\varepsilon}=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,k\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \[u_{i+k,\varepsilon}=\begin{cases} u_{k+1,\varepsilon}u_{i,-\varepsilon}-u_{k,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k$ is odd;}\\ u_{k+1,-\varepsilon}u_{i,\varepsilon}-u_{k,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k$ is even.} \end{cases}\]
  \end{lemma}
  \begin{proof}
    We work by induction on $k$, the cases $k=0,1$ being tautological and reproducing the defining recursion respectively.
    Using the claim for $k$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+k+1,\varepsilon}&=\begin{cases} u_{k+1,\varepsilon}u_{i+1,-\varepsilon}-u_{k,-\varepsilon}u_{i,\varepsilon} & \text{if $k$ is odd;}\\ u_{k+1,-\varepsilon}u_{i+1,\varepsilon}-u_{k,\varepsilon}u_{i,-\varepsilon} & \text{if $k$ is even;} \end{cases}\\
      &=\begin{cases} (u_{2,-\varepsilon}u_{k+1,\varepsilon}-u_{k,-\varepsilon})u_{i,\varepsilon}-u_{k+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k$ is odd;}\\ (u_{2,\varepsilon}u_{k+1,-\varepsilon}-u_{k,\varepsilon})u_{i,-\varepsilon}-u_{k+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k$ is even;} \end{cases}\\
      &=\begin{cases} u_{k+2,-\varepsilon}u_{i,\varepsilon}-u_{k+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k+1$ is even;}\\ u_{k+2,\varepsilon}u_{i,-\varepsilon}-u_{k+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k+1$ is odd;} \end{cases}
    \end{align*}
    which is the claimed recursion for $k+1$.
  \end{proof}

@@

  The Chebyshev polynomials, in particular, allow us to explicitly compute $\phi_{t_i}(\lambda)$ for $\lambda\in\cI$.
  \begin{lemma}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,-}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,+}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,-}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,+}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$ and $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$.
    For $j\ge 0$, using that $(bc-1)u_{2j+1,-}-bu_{2j,-}=u_{2j+3,-}$ and $(bc-1)u_{2j,+}-bu_{2j-1,+}=u_{2j+2,+}$, we have
    \[\phi_{t_{2j+2}(\lambda)}=\phi_{t_2}\phi_{t_{2j}}(\lambda)=(u_{2j+3,-}\lambda_1+u_{2j+2,+}\lambda_2,-u_{2j+2,-}\lambda_1-u_{2j+1,+}\lambda_2).\]
  \end{proof}

>>>>

  \section{Rank Two}
  Fix integers $b,c>0$ and assume $bc\ge 4$.
  Let $B=\left[ \begin{array}{cc} 0 & -b \\ c & 0 \end{array} \right]$.
  Here we may take $\TT_2$ to have vertices labelled $t_\ell$ for $\ell\in\ZZ$, with $t_{2j}$ joined to $t_{2j+1}$ by an edge labelled $1$ and $t_{2j-1}$ joined to $t_{2j}$ by an edge labelled $2$.
  We aim to understand the dominance regions $\cP(\lambda) = \bigcap_{i\in\ZZ} \phi_{t_i}^{-1} \cC_{t_i}(\phi_{t_i} \lambda)$ for $\lambda\in\ZZ^2$.

@@

  It will be useful to have explicit expressions for $\phi_{t_i}(\lambda)$ for $\lambda\in\cI$.
  For this, we define \emph{two-parameter Chebyshev polynomials} $u_{i,\varepsilon}$ for $i\in\ZZ$ and $\varepsilon\in\{\pm\}$ recursively by
  \[u_{0,\varepsilon}=0,\quad u_{1,\varepsilon}=1,\quad u_{i+1,\varepsilon}=\begin{cases} bu_{i,-\varepsilon}-u_{i-1,\varepsilon} & \text{if $\varepsilon=+$;}\\ cu_{i,-\varepsilon}-u_{i-1,\varepsilon} & \text{if $\varepsilon=-$.} \end{cases}\]
  Observe that, by an easy induction, $u_{-i,\varepsilon}=-u_{i,\varepsilon}$.

  The recursion can be written more succinctly as $u_{i+1,\varepsilon}=u_{2,\varepsilon}u_{i,-\varepsilon}-u_{i-1,\varepsilon}$ with the additional initial condition $u_{2,\varepsilon}=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,k\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \begin{equation}
      \label{eq:long Chebyshev recursion}
      u_{i+k,\varepsilon}=\begin{cases} u_{k+1,\varepsilon}u_{i,-\varepsilon}-u_{k,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k$ is odd;}\\ u_{k+1,-\varepsilon}u_{i,\varepsilon}-u_{k,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k$ is even.} \end{cases}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We work by induction on $k$, the cases $k=0,1$ being tautological and reproducing the defining recursion respectively.
    Using the claim for $k>0$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+k+1,\varepsilon}&=\begin{cases} u_{k+1,\varepsilon}u_{i+1,-\varepsilon}-u_{k,-\varepsilon}u_{i,\varepsilon} & \text{if $k$ is odd;}\\ u_{k+1,-\varepsilon}u_{i+1,\varepsilon}-u_{k,\varepsilon}u_{i,-\varepsilon} & \text{if $k$ is even;} \end{cases}\\
      &=\begin{cases} (u_{2,-\varepsilon}u_{k+1,\varepsilon}-u_{k,-\varepsilon})u_{i,\varepsilon}-u_{k+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k$ is odd;}\\ (u_{2,\varepsilon}u_{k+1,-\varepsilon}-u_{k,\varepsilon})u_{i,-\varepsilon}-u_{k+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k$ is even;} \end{cases}\\
      &=\begin{cases} u_{k+2,-\varepsilon}u_{i,\varepsilon}-u_{k+1,-\varepsilon}u_{i-1,\varepsilon} & \text{if $k+1$ is even;}\\ u_{k+2,\varepsilon}u_{i,-\varepsilon}-u_{k+1,\varepsilon}u_{i-1,-\varepsilon} & \text{if $k+1$ is odd;} \end{cases}
    \end{align*}
    which is the claimed recursion for $k+1$.
    These calculations can be reverses to show the result for $k<0$.
  \end{proof}

  \begin{lemma}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{t_{2j}}(\lambda)&=(u_{2j+1,+}\lambda_1+u_{2j,+}\lambda_2,-u_{2j,-}\lambda_1-u_{2j-1,-}\lambda_2);\\
      \phi_{t_{2j+1}}(\lambda)&=(-u_{2j+1,+}\lambda_1-u_{2j,+}\lambda_2,u_{2j+2,-}\lambda_1+u_{2j+1,-}\lambda_2).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    As observed above, we have $\phi_{t_{2j}}=\phi_{t_2}^j$.
    For $\lambda\in\cI$, the action of $\phi_{t_2}$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_{t_2}(\lambda)=(u_{3,-}\lambda_1+u_{2,+}\lambda_2,-u_{2,-}\lambda_1-u_{1,+}\lambda_2),\]
    and so
    \begin{align*}
      \phi_{t_{2j+2}(\lambda)}&=\phi_{t_2}\phi_{t_{2j}}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_1+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_2,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_1-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_2 \big)\\
      &=(u_{2j+3,+}\lambda_1+u_{2j+2,+}\lambda_2, -u_{2j+2,-}\lambda_1-u_{2j+1,-}\lambda_2)
    \end{align*}
    where the last equality uses \eqref{eq:long Chebyshev recursion}.

    Similarly, using $\phi_{t_1}(\lambda)=(-\lambda_1,c\lambda_1+\lambda_2)$ for $\lambda\in\cI$, together with the basic Chebyshev recursion and the equality $\phi_{2j+1}=\phi_{t_1}\phi_{t_2}^j$ gives the given formula for $\phi_{t_{2j+1}}$ from that of $\phi_{t_{2j}}$.
  \end{proof}

@@
